\toolname: file-based localization of numerical perturbations in data analysis pipelines

### 2020-04-13 paper.tex

\toolname identifies the components in a pipeline, at the resolution level of a system process, that produce different
results in different execution conditions. First, a directed
bipartite provenance graph is recorded for each pipeline execution, where nodes
represent application processes and files, and edges represent read and
write file accesses (Figure~\ref{fig:provenance-graph}). Second,
transient files, i.e., files that are either deleted during pipeline
execution or modified by multiple processes, are identified and
disambiguated, resulting in a provenance DAG (Directed Acyclic
Graph) in which file nodes have a single parent (in-degree of 1) (Figure~\ref{fig:provenance-dag}). 
DAGs produced in different conditions are then compared, in
a step-by-step execution that prevents the propagation of differences in the pipeline (Figure~\ref{fig:labeled-dag}).
 The resulting labeled graph identifies the
non-reproducible processes in the pipeline. \note{Greg}{clarify why many processes don't appear in the graph as shown} \tristan{Ali, 
do you pre-process the reprozip trace to remove some processes, or does the figure show all the processes in this trace?}

@@

We randomly selected 20 unprocessed subjects 
from the HCP data release S500 
available in \href{https://db.humanconnectome.org}{the ConnectomDB repository}. \tristan{Ali, I can't find any S500 
data release there, please confirm}
For each subject, available data consisted of 1 or 2 T1w images and 1 or 2 T2w images, with 
$256 \times 320 \times 320$ voxels of size $0.7 \times 0.7 \times 0.7$ mm.
Acquisition protocols and parameters are detailed in~\cite{van2013wu}.

@@

We did not observe any within-OS difference in PreFreeSurfer. In
FreeSurfer, we identified 2 processes leading to within-OS differences due
to the use of pseudo-random numbers: image registration with
\texttt{mri\_segreg}, and cortical surface curvature estimations with
\texttt{mris\_curvature}. Fixing the random seed used in FreeSurfer removed
these differences. \note{Greg}{can you make some statement on the magnitude of these differences?} \tristan{Ali, that's for you.}

@@

To illustrate the magnitude of the differences,
Figure~\ref{fig:fnirt_result} compares \fnirt results in Brain Extraction
for a particular subject. Differences appear to be very substantial, in
particular in the frontal and parietal lobes.\note{Greg}{This is very hard to see from the figure because the
luminosity of the two colormaps are so different - the contrast moving from orange to grayscale is difficult. Perhaps use two colors, green and purple for instance, with the same intensity to see if this is more easily seen?} 
\tristan{Ali, I can't find the script to generate this figure in the repo. Could you fix the colors and contrast as per Greg's suggestion?}
\begin{figure}
  \centering
    \includegraphics[width=\columnwidth]{figures/t2w_alignment.png} 
    \caption{Differences between T2 \fnirt results in PreFreeSurfer's Brain Extraction (CentOS6 vs 
    CentOS7). An animated version of the comparison is available \href{https://github.com/big-data-lab-team/HCP-reproducibility-paper/blob/master/figures/pfs_t2w_alignment.gif}{here}.}
    \label{fig:fnirt_result}
\end{figure}

>>>>

\toolname identifies the components in a pipeline, at the resolution level of a system process, that produce different
results in different execution conditions. First, a directed
bipartite provenance graph is recorded for each pipeline execution, where nodes
represent application processes and files, and edges represent read and
write file accesses (Figure~\ref{fig:provenance-graph}). Second,
transient files, i.e., files that are either deleted during pipeline
execution or modified by multiple processes, are identified and
disambiguated, resulting in a provenance DAG (Directed Acyclic
Graph) in which file nodes have a single parent (in-degree of 1) (Figure~\ref{fig:provenance-dag}). 
DAGs produced in different conditions are then compared, in
a step-by-step execution that prevents the propagation of differences in the pipeline (Figure~\ref{fig:labeled-dag}).
 The resulting labeled graph identifies the
non-reproducible processes in the pipeline. \note{Greg}{clarify why many processes don't appear in the graph as shown} \tristan{Ali, 
do you pre-process the reprozip trace to remove some processes, or does the figure show all the processes in this trace?}
\note{Ali}{It only shows processes that belong to the pipeline (e.g. /usr/local/src/freesurfer/) 
by making a query on the traced DB file. 
We exclude processes like [cp, awk, sed, tee, date\dots].}

@@

We randomly selected 20 unprocessed subjects 
from the HCP data release S500 
available in \href{https://db.humanconnectome.org}{the ConnectomDB repository}. \tristan{Ali, I can't find any S500 
data release there, please confirm.}
\note{Ali}{The Dataset is: \emph{WU-Minn HCP Data - 1200 Subjects} in the project of 
\emph{HCP 1200 Subject Release}.
There is a group name \emph{S500 Release Subjects} that my data is downloaded from.}
For each subject, available data consisted of 1 or 2 T1w images and 1 or 2 T2w images, with 
$256 \times 320 \times 320$ voxels of size $0.7 \times 0.7 \times 0.7$ mm.
Acquisition protocols and parameters are detailed in~\cite{van2013wu}.

@@

We did not observe any within-OS difference in PreFreeSurfer. In
FreeSurfer, we identified 2 processes leading to within-OS differences due
to the use of pseudo-random numbers: image registration with
\texttt{mri\_segreg}, and cortical surface curvature estimations with
\texttt{mris\_curvature}. Fixing the random seed used in FreeSurfer removed
these differences. \note{Greg}{can you make some statement on the magnitude of these differences?} \tristan{Ali, that's for you.}
\note{Ali}{
\texttt{mris\_curvature} produces surfaces like lh.inflated.K and lh.inflated.H, 
I only can see differences using \texttt{mri\_diff} command that it shows:

  Volumes differ in pixel data

  maxdiff   0.18604827 at 130715 0 0 0

for lh.inflated.H and

  Volumes differ in pixel data

  maxdiff   0.85469627 at 130682 0 0 0

for lh.inflated.K

If you know other ways to check magnitude of differences please let me know.
\texttt{mri\_segreg} produces different sum files (ASCII file in which summary statistics are saved), 
but differences belong to meta-data like hostname, date, and SynthSeed. We should remove \texttt{mri\_segreg}
because it doesn't create data differences.
The outputs of both executables are in Consider:/data/asalari/ali-tests/paper\_images/
}

@@

To illustrate the magnitude of the differences,
Figure~\ref{fig:fnirt_result} compares \fnirt results in Brain Extraction
for a particular subject. Differences appear to be very substantial, in
particular in the frontal and parietal lobes.\note{Greg}{This is very hard to see from the figure because the
luminosity of the two colormaps are so different - the contrast moving from orange to grayscale is difficult. Perhaps use two colors, green and purple for instance, with the same intensity to see if this is more easily seen?} 
\tristan{Ali, I can't find the script to generate this figure in the repo. Could you fix the colors and contrast as per Greg's suggestion?}
\note{Ali}{Image is updated.
I made this figure manually using ImageMagick tool and other scripts that you can find them in 
Consider:/data/asalari/ali-tests/paper\_images/pfs\_fnirt\_imgs.}

\begin{figure}
  \centering
    \includegraphics[width=\columnwidth]{figures/t2w_alignment.png} 
    \caption{Differences between T2 \fnirt results in PreFreeSurfer's Brain Extraction (CentOS6 vs 
    CentOS7). An animated version of the comparison is available \href{https://github.com/big-data-lab-team/HCP-reproducibility-paper/blob/master/figures/pfs_t2w_alignment.gif}{here}.}
    \label{fig:fnirt_result}
\end{figure}
